---
title: "Making High-quality tables in R"
author: "Sarah Daly"
date: "April 17, 2019"20
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---
Purpose:  Export 2 tables (.png) in the formate of mean +/- stdev of the raw substrate chemical characteristics and the relevant ratios of these chemical characteristics. Raw substrates also include columns on the qualitiave charactierstics ( Foaming, Date). 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Install libraries
```{r functions,echo=T}
#install.packages("tinytex")
#tinytex::install_tinytex()  # install TinyTeX
library(kableExtra)
library(magrittr)
library(formattable)
library(plyr) #plyr should be loaded before dplyr
library(dplyr)
library(tinytex)
library(rmarkdown)
library(knitr)
library(devtools)
library(summarytools)
library(knitr)
library(kableExtra)
library(pander)
library(formattable)
```
#Setworkding directory and loading data
```{r}

setwd("C:/Users/dalys/OneDrive/Desktop/PHD paper notes/BMP/github/SMY Kinetic Fits")


dt <- read.csv("202000502_growthkinetic_input.csv",header=TRUE, na.strings="NA",stringsAsFactors=FALSE)
head(dt)
sum(is.na(dt))
dim(dt)


#do not differntait between mids
dt <- data.frame(lapply(dt, function(x) {
                  gsub("01/18/18", "1", x)
              }))
dt <- data.frame(lapply(dt, function(x) {
                  gsub("02/18/19", "5", x)
              }))

dt <- data.frame(lapply(dt, function(x) {
                  gsub("10/24/18", "4", x)
              }))
dt <- data.frame(lapply(dt, function(x) {
                  gsub("08/09/18", "3", x)
              }))

dt <- data.frame(lapply(dt, function(x) {
                  gsub("06/05/18", "2", x)
              }))

dt <- data.frame(lapply(dt, function(x) {
                  gsub("05/17/19", "6", x)
              }))

#Convert to numerica
sapply(dt, class)
dt[,11:length(dt)] <- sapply(dt[,11:length(dt)],as.character)
dt[,11:length(dt)] <- sapply(dt[,11:length(dt)],as.numeric)
sapply(dt, class)

```

#Growth Kinetic Table
```{r}

#ONly do final day
dt=filter(dt, Day=='Final')

#HRT crit
cdata <- ddply(dt, c("Substrate", "Date","Location","R.2_adj_avg_Chen"), summarise,
               mean = round(mean(HRTcrit_avg, na.rm=TRUE),digits=0))
HRTcrit=paste0(cdata$mean, "(", round(cdata$R.2_adj_avg_Chen,digits=2), ")")

#um 
cdata <- ddply(dt, c("Substrate", "Date","Location","R.2_adj_avg_Chen"), summarise,
               mean = round(mean(um_avg, na.rm=TRUE),digits=2))
um=paste0(cdata$mean, "(", round(cdata$R.2_adj_avg_Chen,digits=2), ")")
um


#Kchen
cdata <- ddply(dt, c("Substrate", "Date","Location","R.2_adj_avg_Chen"), summarise,
               mean = round(mean(Kchen_avg, na.rm=TRUE),digits=2))
kchen=paste0(cdata$mean, "(", round(cdata$R.2_adj_avg_Chen,digits=2), ")")
kchen

#K1
cdata <- ddply(dt, c("Substrate", "Date","Location","R.2_adj_avg_1st"), summarise,
               mean = round(mean(K1_avg, na.rm=TRUE),digits=2))
k1=paste0(cdata$mean, "(", round(cdata$R.2_adj_avg_1st,digits=2), ")")

#K2
cdata <- ddply(dt, c("Substrate", "Date","Location","R.2_adj_avg_2nd"), summarise,
               mean = round(mean(K2_avg, na.rm=TRUE),digits=2))
k2=paste0(cdata$mean, "(", round(cdata$R.2_adj_avg_2nd,digits=2), ")")

#Kt
cdata <- ddply(dt, c("Substrate", "Date","Location"), summarise,
               mean = round(mean(Kt_avg, na.rm=TRUE),digits=2))
kt=paste0(cdata$mean)

#T80
cdata <- ddply(dt, c("Substrate", "Date","Location"), summarise,
               mean = round(mean(T80_avg, na.rm=TRUE),digits=0))
T80=paste0(cdata$mean)

#T90
cdata <- ddply(dt, c("Substrate", "Date","Location"), summarise,
               mean = round(mean(T90_avg, na.rm=TRUE),digits=0))
T90=paste0(cdata$mean)

#lag
cdata <- ddply(dt, c("Substrate", "Date","Location","R.2_adj_avg_Gompertz"), summarise,
               mean = round(mean(lag_avg, na.rm=TRUE),digits=0))
lag=paste0(cdata$mean, "(", round(cdata$R.2_adj_avg_Gompertz,digits=2), ")")
lag

#K1 mod
cdata <- ddply(dt, c("Substrate", "Date","Location","R.2_adj_avg_1stmod"), summarise,
               mean = round(mean(K1_avg.1, na.rm=TRUE),digits=2))
k1mod=paste0(cdata$mean, "(", round(cdata$R.2_adj_avg_1stmod,digits=2), ")")

#K2 mod
cdata <- ddply(dt, c("Substrate", "Date","Location","R.2_adj_avg_1stmod"), summarise,
               mean = round(mean(K2_avg.1, na.rm=TRUE),digits=2))
k2mod=paste0(cdata$mean, "(", round(cdata$R.2_adj_avg_1stmod,digits=2), ")")
k2mod




Location=as.character(cdata$Location)
Date=as.character(cdata$Date)
Substrate=as.character(cdata$Substrate)

new<-cbind(Substrate,Date, Location, HRTcrit, um, kchen, k1,k2,k1mod, k2mod, lag,kt,T80,T90)
new

#<br/> html for new line
  names=c("Substrate Composition     ","Date","Location","HRT<sub>critical</sub> <br/> day","&#956m <br/> day<sup>-1</sup>","k<sub>Chen</sub> <br/>  day<sup>-1</sup>","k<sub>first</sub> <br/>  day<sup>-1</sup>","k<sub>second</sub> <br/>  day<sup>-1</sup>","k<sub>1 modifed first order</sub> <br/> day<sup>-1</sup>","k<sub>2 modified first order</sub> <br/>  day<sup>-1</sup>","lag <br/> day", "k<sub>t</sub> <br/> day<sup>-1</sup>","T<sub>80</sub> <br/> day","T<sub>90</sub> <br/> day")
  

x=kable_styling(kable(new[order(Substrate),], format='html', col.names=names, escape = FALSE,table.attr = "style = \"color: black;\""),"striped","scale_down")  
kableExtra::row_spec(x, 0, extra_css = "border-bottom: 1px solid")
x
 
  save_kable(x, file='Growth Kinetics BMP.png')
  
  #Save as soft text
write.table(new, file = "Growth Kinetics BMP.txt", sep = ",", quote = FALSE, row.names = F)
#save as pdf
save_kable(x, file='Growth Kinetics BMP.pdf')
#Save as png
save_kable(x, file='Growth Kinetics BMP.png')



                       

```


#Methane production based on mL CH4/g vS added
```{r}

#ONly do final day
dt=filter(dt, Day=='Final')

#Chen & Hashimoto
cdata <- ddply(dt, c("Substrate", "Date","Location","B_R.2_adj_avg_pred_Chen"), summarise,
               mean = round(mean(B_pred_avg_Chen, na.rm=TRUE),digits=2))
Predicted_Chen=paste0(cdata$mean, "(", round(cdata$B_R.2_adj_avg_pred_Chen,digits=2), ")")


#Gompertz predicted
cdata <- ddply(dt, c("Substrate", "Date","Location","B_R.2_adj_pred_avg_Gompertz"), summarise,
               mean = round(mean(as.numeric(B_pred_avg_Gompertz), na.rm=TRUE),digits=2))
Predicted_Gompertz=paste0(cdata$mean, "(", round(as.numeric(cdata$B_R.2_adj_pred_avg_Gompertz),digits=2), ")")

#First predicted
cdata <- ddply(dt, c("Substrate", "Date","Location","B_R.2_adj_avg_pred_1st"), summarise,
               mean = round(mean(B_pred_avg_1st, na.rm=TRUE),digits=2))
Predicted_1st=paste0(cdata$mean, "(", round(cdata$B_R.2_adj_avg_pred_1st,digits=2), ")")

#Second predicted
cdata <- ddply(dt, c("Substrate", "Date","Location","B_R.2_adj_avg_B_avg_pred_2nd"), summarise,
               mean = round(mean(B_pred_avg_2nd, na.rm=TRUE),digits=2))
Predicted_2nd=paste0(cdata$mean, "(", round(cdata$B_R.2_adj_avg_B_avg_pred_2nd,digits=2), ")")

#First modified predicted
cdata <- ddply(dt, c("Substrate", "Date","Location","B_R.2_adj_avg_pred_1stmod"), summarise,
               mean = round(mean(as.numeric(B_pred_avg_1stmod), na.rm=TRUE),digits=2))
Predicted_1stmod=paste0(cdata$mean, "(", round(as.numeric(cdata$B_R.2_adj_avg_pred_1stmod),digits=2), ")")

#Methane Yield
cdata <- ddply(dt, c("Substrate", "Date","Location"), summarise,
               mean = round(mean(CH4_Yield_VS_added, na.rm=TRUE),digits=2))
Methane_Yield=paste0(cdata$mean)


Location=as.character(cdata$Location)
Date=as.character(cdata$Date)
Substrate=as.character(cdata$Substrate)
#Methane_Yield=as.character(round(cdata$CH4_Yield_VS_added,digits=0))

new<-cbind(Substrate,Date, Location, Methane_Yield,Predicted_Chen,Predicted_Gompertz,Predicted_1st,Predicted_2nd,Predicted_1stmod)
new

names =c("Substrate Composition","Date","Location","BMP <br/> mL CH<sub>4</sub> gVSadded<sup>-1</sup>","Predicted Chen & Hashimoto BMP <br/> mL CH<sub>4</sub> gVSadded<sup>-1</sup>","Predicted Gompertz BMP <br/> mL CH<sub>4</sub> gVSadded<sup>-1</sup>","Predicted First Order BMP <br/> mL CH<sub>4</sub> gVSadded<sup>-1</sup>", "Predicted Second Order BMP <br/> mL CH<sub>4</sub> gVSadded<sup>-1</sup>","Predicted Modified First Order BMP <br/> mL CH<sub>4</sub> gVSadded<sup>-1</sup>")

x=kable_styling(kable(new[order(Substrate),], format='html', col.names=names, escape = FALSE,table.attr = "style = \"color: black;\""),"striped","scale_down")  
kableExtra::row_spec(x, 0, extra_css = "border-bottom: 1px solid")
x


kable_styling(kable(new[order(Substrate),], format='html', col.names=names, escape = FALSE,table.attr = "style = \"color: black;\""),bootstrap_options = c("striped", "hover", "condensed","bordered"),full_width = F)
kableExtra::row_spec(x, 0, extra_css = "border-bottom: 1px solid")
#kableExtra::as_image(width = 8,density=800)


save_kable(x, file='BMP v Predicted BMP.png')

#Save as soft text
write.table(new, file = "BMP v Predicted BMP.txt", sep = ",", quote = FALSE, row.names = F)
#save as pdf
save_kable(x, file='BMP v Predicted BMP.pdf')
#Save as png
save_kable(x, file=' BMP v Predicted BMP.png')


```
  

  
  
#CH4 value for each treatment

```{r}
#ONly do final day
dt=filter(dt, Day=='Final')


#CH4 g TCOD added
cdata <- ddply(dt, c("Substrate", "Date","Day"), summarise,
               mean = round(mean(CH4_Yield_TCOD_added, na.rm=TRUE),digits=4),
              sd = round(sd(CH4_Yield_TCOD_added, na.rm=TRUE),digits=4)
              
)
CH4TCOD=paste0(cdata$mean, " ? ", cdata$sd)

#CH4 g VS added
cdata <- ddply(dt, c("Substrate", "Date","Day"), summarise,
               mean = round(mean(CH4_Yield_VS_added, na.rm=TRUE),digits=4),
              sd = round(sd(CH4_Yield_VS_added, na.rm=TRUE),digits=4)
              
)
CH4VSadded=paste0(cdata$mean, " ? ", cdata$sd)


Date=as.character(cdata$Date)
Day=as.character(cdata$Day)
Substrate=as.character(cdata$Substrate)

new<-cbind(Date, Day, Substrate, CH4VSadded, CH4TCOD)
new

#<br/> html for new line
  names=c("Test No.","Day","Treatment", "CH4 Yield g VS added","CH4 g TCOD added")

    #options(knitr.table.format = "html")
x=kable_styling(kable(new, format='html', col.names=names, escape = FALSE,table.attr = "style = \"color: black;\""),"striped","scale_down")
x

#Save as soft text
write.table(new, file = "CH4 yield.txt", sep = ",", quote = FALSE, row.names = F)
#save as pdf
save_kable(x, file=' CH4 yield.pdf')
#Save as png
save_kable(x, file=' CH4 yield.png')





```


#Generate a table of pearson correlations
http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r
https://statistics.laerd.com/statistical-guides/pearson-correlation-coefficient-statistical-guide.php
https://bookdown.org/ndphillips/YaRrr/correlation-cor-test.html

```{r}
library(magicfor)
#Filter out Day 0
dt=filter(dt, Day=='Zero')
x<-cbind(dt$H2S_yield_gVSadded,dt$CONDUCTIVITY,dt$SCOD,dt$SCOD.TCOD,dt$TKN,dt$TN,dt$INORGN,dt$TVFA,dt$TALK,dt$TANNIC, dt$TAN,dt$FE,dt$SULFATE,dt$TCOD.SULFATE,dt$pH,dt$CU,dt$NI,dt$TP, dt$OP, dt$OP.TP,dt$TVFA.TALK,dt$TAN.TKN,dt$TCOD.TKN,dt$FE.II..S, dt$Fe.II..TP,dt$TCOD.TN.TP)


#Replace 0 with small value 
x[x==0]<-0.00001
x_trans=log(na.omit(x))
#x_trans=na.omit(x)
num=dim(x)[2]
i=1
opp1<-list()
opp2<-list()
opp3<-list()
while (i<=num){
op1<-c() #Initialize a clean list
op2<-c()
op3<-c()
#Check normality of data
t.lm=lm(x_trans[,1]~x_trans[,i])
norm_check=shapiro.test(t.lm$residuals) #If P-value <0.05 than the data is not normally distriuted
print(norm_check)
#Use pearson for normal, spearman for non-normal
if (norm_check[2]<0.05){coeff=cor(x_trans[,1], x_trans[,i], method = "spearman")
method="spearman"}
else {
coeff=cor(x_trans[,1], x_trans[,i], method = "pearson")
method="pearson"}
pval=cor.test(x_trans[,1],x_trans[,i], method=method)$p.value
print(coeff)
op1<-coeff  #Put results in a clean list
opp1[[i]]<-op1 #Create a list of lists
op2<-method
opp2[[i]]<-op2
op3<-pval
opp3[[i]]<-op3
i=i+1
}


opp1=do.call("rbind", opp1)
opp2=do.call("rbind", opp2)
opp3=do.call("rbind", opp3)

pars<-c("H2SgVSadded","Conductivity","SCOD","SCOD.TCOD","TKN","TN","INORGN","TVFA", "TALK","TANNIC","TAN","FE","Sulfate","TCOD.Sulfate","pH","CU","NI","TP","OP","OP.TP","TVFA.TALK","TAN.TKN","TCOD.TKN","FE(II).S","Fe(II).TP","TCOD.TN.TP")

p<-cbind(pars,opp1,opp2,opp3)

cormatrix<-as.data.frame(p)
names(cormatrix)<-c("Parameter","Correlation Coefficient","Test","pval")
cormatrix



#cor(sqrt(na.omit(x)))
#comatrix=cor(log(na.omit(x)))
#comatrix=cor(x_trans)
#pars<-c("H2SgVSadded","Conductivity","SCOD","SCOD.TCOD","TKN","TN","INORGN","TVFA", "TALK","TANNIC","TAN","FE","Sulfate","TCOD.Sulfate","pH","CU","NI","TP","OP","OP.TP","TVFA.TALK","TAN.TKN","TCOD.TKN","FE(II).S","Fe(II).TP","TCOD.TN.TP")
#corrframe=cbind(pars,comatrix[,1])
#corrframe

#<br/> html for new line
#names=c("Parameter.","Coefficient")

    #options(knitr.table.format = "html")
x=kable_styling(kable(cormatrix, format='html', escape = FALSE,table.attr = "style = \"color: black;\""),"striped","scale_down")
x

#Save as soft text
write.table(cormatrix, file = "Correlation H2S yield.txt", sep = ",", quote = FALSE, row.names = F)
#save as pdf
save_kable(x, file=' Correlation H2S yield.pdf')
#Save as png
save_kable(x, file='Correlation H2S yield.png')
```
